Adaptive broadcast tree for multicores

### 2013-07-01 paper.tex
Improved Introduction

Distributed algorithms are increasingly used to overcome scalability
challenges on multicore machines. Examples are
databases~\cite{Salomie2011, Wiesmann2000} and operating
systems~\cite{fos:osr09, tornado:osdi99, barrelfish:sosp09}.

% What do you want to enable?
We want to investigate how to achieve better performance of atomic
broadcast across a wide range of multicore machines without manually
tuning the implementation to concrete machines.

% What problem are you solving, and why is it hard?
The problem is complexity and diversity of modern multicore
machines. Multicores come in many flavors depending on vendor and
generation of the machine. All of these machines have fundamentally
different characteristics such as number of nodes, NUMA
characteristics, interconnect topologies, propagation time on
interconnect and use of shared caches. These characteristics are hard~\cite{Cavage2013}
to understand, but important to consider for application performance.

% New ideas
Instead of manually tuning algorithms to characteristics of individual
machines, we automatically configure distributed algorithm
implementations based on a machine model. We represent this model as an
annotated graph. One question we are looking at is which information
should actually be part of this machine model.

% How will you go about it?
We base the configuration of algorithms on \emph{overlay networks}, which we
automatically configure. We execute our configuration stage on a model
representing the concrete machine the algorithm is intended to run
on. %
Our idea is to then explore different approaches for finding overlay
networks based on the network model graph. Approaches we are currently
investigating are: minimum spanning tree algorithms, binary trees,
clustering techniques as well as treating the problem
hierarchically. Assuming the overlay network, we then need to decide
on every node which message to send first. This is a \emph{scheduling}
problem.

>>>>

Distributed algorithms are increasingly used to overcome scalability
challenges of multicore machines. Examples are
databases~\cite{Salomie2011, Wiesmann2000} and operating
systems~\cite{fos:osr09, tornado:osdi99, barrelfish:sosp09}.

% What do you want to enable?
We want to investigate how to achieve better performance of these distributed
algorithms across a wide range of multicore machines without manually
tuning the implementation to characteristics of individual machines.

% What problem are you solving, and why is it hard?
The problem we are tackling is an increase in complexity and diversity
of modern multicore machines. Multicores come in many flavors
depending on vendor and product line of the machine. They also change
heavily over time. Consequently, characteristics of such machines are
fundamentally different. Examples are the number of cores, NUMA
configurations, interconnect topologies, propagation time on the
interconnect and availability of shared resources such as
caches. These characteristics are hard~\cite{Cavage2013} to
understand, but important to consider for application performance.

% New ideas
Instead of manually tuning algorithms to characteristics of individual
machines, we automatically configure distributed algorithm based on a
machine model. This is populate with online measurements to express
detailed information about machine characteristics. We then use
this model to predict algorithm performance for the specific machine
and select a good implementation automatically and dynamically at
run-time.

% How will you go about it?
As an example, we investigate atomic broadcasts for multicore
machines. The space of possible implementations is large reaching from
shared memory implementations to message based topologies for
different topologies of overlay networks. Such topologies are: minimum
spanning trees, binary trees, rings as well as clustering-based
approaches. Assuming the overlay network, we then need to decide for
every node the order in which to send messages. This is a
\emph{scheduling} problem.
